---
title: "MH3511 Team Project"
output: html_notebook
---

# Data Science Fields Salary Categorization Dataset

## Overview
This dataset aims to shed light on the salary trends in the field of Data Science for the years 2021 to 2023. With a focus on various aspects of employment, including work experience, job designation, and company locations, this dataset provides valuable insights into salary distributions within the industry.

## Load packages
```{r}
install.packages("dplyr")
library(dplyr)
```

## Import Dataset
```{r}
# Change Directory to Location of Dataset
df <- read.csv("/Users/nicoleang/Downloads/Data_Science_Fields_Salary_Categorization.csv")

# Display the first few rows
head(df)
summary(df)
```

## Data Cleaning
```{r}
# Check for missing values in each column
colSums(is.na(df))
```

```{r}
# Check for any duplicate rows
sum(duplicated(df))
```

```{r}
# Dropping of irrelevant columns (Employee Location - Not used in the analysis)
df <- subset(df, select = -c(1, Employee_Location))
```

```{r}
# Dropping of observations in Employment_Status that are not 'FT'

# Check for unique values under Employment_Status
table(df$Employment_Status)
# Remove observations where Employment_Status is 'CT', 'PT', 'FL'
df <- subset(df, !Employment_Status %in% c('CT', 'PT', 'FL'))
```

```{r}
# Convert Salary from Rupees to SGD
# Convert Salary from character string to numeric values
df$Salary_In_Rupees <- as.numeric(gsub(",", "", df$Salary_In_Rupees))
# Currency Conversion
df$Salary_In_Rupees <- df$Salary_In_Rupees * 0.016
# Rename Salary_In_Rupees to Salary_In_SGD
colnames(df)[colnames(df) == "Salary_In_Rupees"] <- "Salary_In_SGD"
head(df)
```


```{r}
table(df$Designation)
# Filter observations to only include Designation as 'Data Scientist', 'Data Engineer', 'Data Analyst', 'Machine Learning Engineer'
df <- df[df$Designation %in% c("Data Scientist", "Data Engineer", "Data Analyst", "Machine Learning Engineer"), ]
```

## Summary Statistics

### Working_Year
```{r}
year_count <- table(df$Working_Year)
barplot(year_count, 
        main = "Working Year", 
        xlab = "Working_Year", 
        ylab = "Frequency", 
        col = "grey")
```
### Designation
```{r}
# Split "Designation" column into lists
split_desgn <- strsplit(df$Designation, ",")
# Remove extra spaces
split_desgn <- lapply(split_desgn, trimws)
# Create new data frame with individual categories
df.desgn <- data.frame(Designation = unlist(split_desgn))
# Plot bar plot
barplot(table(df.desgn$Designation), main="Designation")
```
### Experience
```{r}
table(df$Experience)
# Split "Experience" column into lists
split_exp <- strsplit(df$Experience, ",")
# Remove extra spaces
split_exp <- lapply(split_exp, trimws)
# Create new data frame with individual categories
df.exp <- data.frame(Experience = unlist(split_exp))
# Plot bar plot
barplot(table(df.exp$Experience), main="Experience Level")

# Check number of observations with Experience == 'EX'
sum(df.exp$Experience == "EX")
# Remove observations with 'EX' as Experience
df <- df[df$Experience != "EX", ]
# Split "Experience" column into lists
split_exp <- strsplit(df$Experience, ",")
# Remove extra spaces
split_exp <- lapply(split_exp, trimws)
# Create new data frame with individual categories
df.exp <- data.frame(Experience = unlist(split_exp))
# Plot new bar plot
barplot(table(df.exp$Experience), main="Experience Level")
```
### Company_Location
```{r}
# Split "Company_Location" column into lists
split_loc <- strsplit(df$Company_Location, ",")
# Remove extra spaces
split_loc <- lapply(split_loc, trimws)
# Create new data frame with individual categories
df.loc <- data.frame(Company_Location = unlist(split_loc))

# Create a table with the count of each Company_Location
location_count <- table(df.loc$Company_Location)
# Filter for locations with count greater than 10
filtered_location_count <- location_count[location_count > 10]
# Print the filtered values and counts
print(filtered_location_count)

# Plot bar plot for filtered values
barplot(filtered_location_count, main="Company Location (count > 10)")
```
```{r}
# Filter for Top 6 countries
top_six <- subset(df, Company_Location %in% c("US", "GB", "CA", "DE", "IN", "FR"))

# Order countries by median salary
country_stats <- top_six %>%
  group_by(Company_Location) %>%
  summarize(
    median_salary = median(Salary_In_SGD),
    mean_salary = mean(Salary_In_SGD),
    count = n()
  ) %>%
  arrange(desc(median_salary))

country_stats
```
### Company_Size
```{r}
# Split "Company_Size" column into lists
split_csize <- strsplit(df$Company_Size, ",")
# Remove extra spaces
split_csize <- lapply(split_csize, trimws)
# Create new data frame with individual categories
df.csize <- data.frame(Company_Size = unlist(split_csize))
# Plot bar plot
barplot(table(df.csize$Company_Size), main="Company Size")
```
### Remote_Working_Ratio
```{r}
ratio_count <- table(df$Remote_Working_Ratio)
barplot(ratio_count, 
        main = "Remote Working Ratio", 
        xlab = "Remote_Working_Ratio", 
        ylab = "Frequency", 
        col = "grey")
```
### Salary_In_SGD
```{r}
# Plot overall distribution of variable Salary_In_SGD
hist(df$Salary_In_SGD, xlab="Salary_In_SGD", main="Histogram of Salary In SGD")
boxplot(df$Salary_In_SGD, main="Box Plot of Salary In SGD", ylab="Salary_In_SGD")

# Remove outliers, Salary_In_SGD > 300000
df <- df[df$Salary_In_SGD <= 300000, ]
# Plot new overall distribution of variable Salary_In_SGD
hist(df$Salary_In_SGD, xlab="Salary_In_SGD", main="Histogram of Salary In SGD")
boxplot(df$Salary_In_SGD, main="Box Plot of Salary In SGD", ylab="Salary_In_SGD")
summary(df)
```
## Save cleaned dataset as .csv file
```{r}
write.csv(df, "DataScienceSalaries.csv", row.names = FALSE)
```